Automatically Splitting a Staged Lambda Calculus

### 2015-07-08 paper/relatedwork.tex
small changes to related work

\paragraph{Frequency-reduction via pass separation.}
Perhaps the closest work to ours is the algorithm for data
specialization given by Knoblock and Ruf~\cite{knoblock96}, which
also seeks to statically split an explicitly staged program into two
stages.  However, they only consider a simple first-order language and
straight-line programs. Their work also treats all computations
guarded by second-stage conditionals as second-stage computations,
which would prevent optimization (via splitting) of programs such as
quickselect.

Automatic pass separation has also been a topic of interest in
programming systems for computer graphics, where, to achieve
high performance, programs are typically manually separated into
components by frequency of execution (Indeed, separation into a pre-determined
set of per-frequency modules is \emph{required} by the execution model
of modern graphics architectures.)  The software engineering
challenges of modifying multiple per-stage programs have led to
suggestions of writing graphics programs in an explicitly-staged
programming language\,\cite{Proudfoot:2001,Foley:2011,He:2014}, and
deferring the task of pass separation to the compiler. However, all
prior efforts, like that of Knoblock and Ruf\,\cite{knoblock96} have
been limited to simple, imperative languages. We are not aware of
prior work that presents a splitting algorithm in the context of
higher-order functional languages.

>>>>

\paragraph{Stage Splitting.}

Algorithms similar to ours have appeared in the partial evaluation community,
where they are used as a preprocessing step to partial evaluation. 
For instance, \cite{Mogensen89a} and \cite{DeNiel91} defined splitting algorithms
for programs consistening of a set of first-order recursive functions.
In addition to the lack of first-class functions, 
the binding time specificiations used by these systems are more complicated and less expressive than the type system of \lang.
The question of binding time analysis for a higher-order language was visited in \cite{Mogensen89b},
however the implications for splitting were not explicitly considered.

Automatic pass separation has also been a topic of interest in
programming systems for computer graphics, where to achieve
high performance, programs are typically manually separated into
components by frequency of execution. (Indeed, separation into a pre-determined
set of per-frequency modules is \emph{required} by the execution model
of modern graphics architectures.)  
The software engineering challenges of modifying programs in the split form (see \cite{Foley:2011}) have led to
suggestions of differing the task of splitting to the compiler.
Early work (\cite{knoblock96}) sought to determine the staging of programs automatically through use of binding time analysis,
however later work (\cite{Proudfoot:2001,Foley:2011,He:2014}) has focused on giving programmers explicit control of the staging through 
programming language features. 
All of these efforts in graphics have been limited to simple, imperative languages.
